<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{ERDDAP vignette}
-->

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>", 
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  cache = TRUE
)
```

ERDDAP vignette
======

ERDDAP gives a simple, consistent way to download subsets of gridded and tabular scientific datasets in common file formats and make graphs and maps. This ERDDAP installation has oceanographic data (for example, data from satellites and buoys).

Find out more about ERDDAP at [http://coastwatch.pfeg.noaa.gov/erddap/index.html](http://coastwatch.pfeg.noaa.gov/erddap/index.html).

## Load rnoaa

```{r}
library('rnoaa')
```

## Passing the datasetid without fields gives all columns back

```{r one}
erddap_table('erdCalCOFIfshsiz')
```

## Pass time constraints

```{r two}
erddap_table('erdCalCOFIfshsiz', 'time>=2001-07-07', 'time<=2001-07-08')
```

## Pass in fields (i.e., columns to retrieve) & time constraints

```{r three}
erddap_table('erdCalCOFIfshsiz', fields=c('longitude','latitude','fish_size','itis_tsn'), 'time>=2001-07-07','time<=2001-07-10')
```

```{r}
erddap_table('erdCinpKfmBT', fields=c('latitude','longitude','Aplysia_californica_Mean_Density','Muricea_californica_Mean_Density'), 'time>=2007-06-24','time<=2007-07-01')
```

## An example workflow

Search for data

```{r}
(out <- erddap_search(query='size'))
```

Or, list datasets

```{r}
tdp <- erddap_datasets('table')
```

Using a datasetid, search for information on a datasetid

```{r}
id <- as.character(tdp$Dataset.ID[8])
erddap_info(id)$variables
```

Get data from the dataset

```{r}
erddap_table(id, fields = c('latitude','longitude'))
```

## Time constraint

Limit by time with date only

```{r}
erddap_table(id, fields = c('latitude','longitude','time'), 'time>=2001-07-14', 'time<=2001-09-14')
```

## Use distinct parameter

```{r}
erddap_table('erdCalCOFIfshsiz', fields=c('longitude','latitude','fish_size','itis_tsn'), 'time>=2001-07-07','time<=2001-07-10', distinct=TRUE)
```

## The units parameter

In this example, values are the same, but sometimes they can be different given the units value passed

```{r}
erddap_table('erdCinpKfmT', fields=c('longitude','latitude','time','temperature'), 'time>=2007-09-19', 'time<=2007-09-21', units='udunits')
```

```{r}
erddap_table('erdCinpKfmT', fields=c('longitude','latitude','time','temperature'), 'time>=2007-09-19', 'time<=2007-09-21', units='ucum')
```

## The orderby parameter

```{r}
erddap_table('erdCinpKfmT', fields=c('longitude','latitude','time','temperature'), 'time>=2007-09-19', 'time<=2007-09-21', orderby='temperature')
```

## The orderbymax parameter

```{r}
erddap_table('erdCinpKfmT', fields=c('longitude','latitude','time','temperature'), 'time>=2007-09-19', 'time<=2007-09-21', orderbymax='temperature')
```

## The orderbymin parameter

```{r}
erddap_table('erdCinpKfmT', fields=c('longitude','latitude','time','temperature'), 'time>=2007-09-19', 'time<=2007-09-21', orderbymin='temperature')
```

## The orderbyminmax parameter

```{r}
erddap_table('erdCinpKfmT', fields=c('longitude','latitude','time','temperature'), 'time>=2007-09-19', 'time<=2007-09-21', orderbyminmax='temperature')
```

## The orderbymin parameter with multiple values

```{r}
erddap_table('erdCinpKfmT', fields=c('longitude','latitude','time','depth','temperature'), 'time>=2007-06-10', 'time<=2007-09-21', orderbymax=c('depth','temperature'))
```

## Spatial delimitation

```{r}
erddap_table('erdCalCOFIfshsiz', fields = c('latitude','longitude','scientific_name'), 'latitude>=34.8', 'latitude<=35', 'longitude>=-125', 'longitude<=-124')
```

## Integrate with taxize

```{r}
out <- erddap_table('erdCalCOFIfshsiz', fields = c('latitude','longitude','scientific_name','itis_tsn'))
tsns <- unique(out$itis_tsn[1:100])
library("taxize")
classif <- classification(tsns, db = "itis")
head(rbind(classif)); tail(rbind(classif))
```
