<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{HOMR metadata}
-->

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>", 
  warning = FALSE, 
  message = FALSE,
  cache = TRUE
)
```


HOMR metadata
======

`HOMR` (Historical Observing Metadata Repository) provides climate station metadata. It's a NOAA service.

Find out more about HOMR at [http://www.ncdc.noaa.gov/homr/](http://www.ncdc.noaa.gov/homr/) and the HOMR API at [http://www.ncdc.noaa.gov/homr/api](http://www.ncdc.noaa.gov/homr/api).

## Load rnoaa

```{r load}
library('rnoaa')
```

## Search by station identifier

You can do this in various ways. Using the `qid` parameter (stands or qualified ID, as far as I know), you can search by suffix (e.g., `046742`), or both separated by a colon (e.g., `COOP:046742`). 

By station suffix

```{r}
res <- homr(qid = ':046742')
names(res)
names(res[['20002078']])
res$`20002078`[1:3]
```

By both

```{r}
res <- homr(qid = 'COOP:046742')
names(res)
names(res[['20002078']])
res$`20002078`[1:5]
```

## Search by station parameter

You can also search by station identifier, which is different from the `qid` above. 

```{r}
res <- homr(station=20002078)
names(res)
names(res[['20002078']])
res$`20002078`[4:6]
```

## Search by state and county

By state

```{r}
res <- homr(state='DE', begindate='2005-01-01', enddate='2005-02-01')
names(res)
```

By country

```{r}
res <- homr(country='GHANA', begindate='2005-01-01', enddate='2005-02-01')
library("plyr")
ldply(res, function(x) x$location$latlon)
```

By state and county

```{r}
res <- homr(state='NC', county='BUNCOMBE', headersOnly = TRUE)
head( ldply(res, "[[", "head") )
```

## Get header information only

```{r}
res <- homr(headersOnly=TRUE, state='DE')
head( ldply(res, "[[", "head") )
```

## Data definitions

The data returned is the same format for all, so a separate function is provided to get metadata. The function `homr_definitions()` does query the HOMR API, so does get updated metadata - i.e., it's not a static dataset stored locally. 

```{r}
head( homr_definitions() )
```
