<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{rnoaa vignette}
-->

rnoaa vignette
======

### About the package

`rnoaa` is an R wrapper for the NOAA API.

********************

### Install `rnoaa`

Install and load `rnoaa` into the R session.

```{r install, eval=FALSE}
install.packages("devtools")
library(devtools)
install_github("rnoaa", "ropensci", ref="newapi")
```

```{r load, message=FALSE, warning=FALSE}
library(rnoaa)
library(plyr)
```

#### Get info on a station by specifying a datasetid, locationid, and stationid

```{r two, message=FALSE, warning=FALSE, cache=FALSE}
noaa_stations(datasetid='GHCND', locationid='FIPS:12017', stationid='GHCND:USC00084289')
```

#### Search for data and get a data.frame or list

```{r three, message=FALSE, warning=FALSE, cache=FALSE}
out <- noaa(datasetid='NORMAL_DLY', stationid='GHCND:USW00014895', datatypeid='dly-tmax-normal')
```

See a data.frame

```{r four, message=FALSE, warning=FALSE, cache=FALSE}
out$data
```

#### Plot data, super simple, but it's a start

```{r six, message=FALSE, warning=FALSE, cache=FALSE}
out <- noaa(datasetid='NORMAL_DLY', stationid='GHCND:USW00014895', datatypeid='dly-tmax-normal')
noaa_plot(out)
```

### More on plotting

#### Example 1

Search for data first, then plot

```{r}
out <- noaa(datasetid='GHCND', stationid='GHCND:USW00014895', datatypeid='PRCP', startdate = '2010-05-01', enddate = '2010-10-31', limit=500)
```

Default plot

```{r}
noaa_plot(out)
```

Create 14 day breaks

```{r}
noaa_plot(out, breaks="14 days")
```

One month breaks

```{r}
noaa_plot(out, breaks="1 month", dateformat="%d/%m")
```

#### Example 2

Search for data

```{r}
out2 <- noaa(datasetid='GHCND', stationid='GHCND:USW00014895', datatypeid='PRCP', startdate = '2010-05-01', enddate = '2010-05-03', limit=100)
```

Make a plot, with 6 hour breaks, and date format with only hour

```{r}
noaa_plot(out2, breaks="6 hours", dateformat="%H")
```

#### Combine many calls to noaa function

Search for two sets of data

```{r}
out1 <- noaa(datasetid='GHCND', stationid='GHCND:USW00014895', datatypeid='PRCP', startdate = '2010-03-01', enddate = '2010-05-31', limit=500)

out2 <- noaa(datasetid='GHCND', stationid='GHCND:USW00014895', datatypeid='PRCP', startdate = '2010-09-01', enddate = '2010-10-31', limit=500)
```

Then combine with a call to `noaa_combine`

```{r}
df <- noaa_combine(out1, out2)
head(df[[1]]); tail(df[[1]])
```

Then plot - the default passing in the combined plot plots the data together. In this case it looks kind of weird since a straight line combines two distant dates.

```{r}
noaa_plot(df)
```

But we can pass in each separately, which uses `facet_wrap` in `ggplot2` to plot each set of data in its own panel.

```{r}
noaa_plot(out1, out2, breaks="45 days")
```