<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Sea ice vignette}
%\VignetteEncoding{UTF-8}
-->

Sea ice vignette
======

Get sea ice data at [ftp://sidads.colorado.edu/DATASETS/NOAA/G02135/shapefiles](ftp://sidads.colorado.edu/DATASETS/NOAA/G02135/shapefiles).

********************

### Install `rnoaa`

Install and load `rnoaa` into the R session.

If you're on Windows, you may have to install Rtools. Run `devtools::has_devel()`. If you get a `TRUE`, you're okay. If not, [install Rtools](http://cran.r-project.org/bin/windows/Rtools/).

```{r install, eval=FALSE}
install.packages("devtools")
devtools::install_github("ropensci/rnoaa")
```

```{r load, message=FALSE, warning=FALSE}
library('rnoaa')
library('plyr')
```

### Look at data.frame's for a series of years for Feb, South pole

```{r message=FALSE, warning=FALSE}
urls <- sapply(seq(1979,1990,1), function(x) seaiceeurls(yr=x, mo='Feb', pole='S'))
out <- lapply(urls, seaice)
head(out[[1]])
```

### Map a single year/month/pole combo

```{r}
urls <- seaiceeurls(mo='Apr', pole='N', yr=1990)
out <- seaice(urls)
library('ggplot2')
ggplot(out, aes(long, lat, group=group)) +
   geom_polygon(fill="steelblue") +
   theme_ice()
```

### Map all years for April only for North pole

```{r message=FALSE, warning=FALSE}
urls <- seaiceeurls(mo='Apr', pole='N')[1:10]
out <- lapply(urls, seaice)
names(out) <- seq(1979,1988,1)
df <- ldply(out)
library('ggplot2')
ggplot(df, aes(long, lat, group=group)) +
  geom_polygon(fill="steelblue") +
  theme_ice() +
  facet_wrap(~ .id)
```
