<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{SWDI vignette}
-->

SWDI vignette
======

`SWDI` is the Severe Weather Data Inventory. SWDI is an

> integrated database of severe weather records for the United States. The records in SWDI come from a variety of sources in the NCDC archive. SWDI provides the ability to search through all of these data to find records covering a particular time period and geographic region, and to download the results of your search in a variety of formats. The formats currently supported are Shapefile (for GIS), KMZ (for Google Earth), CSV (comma-separated), and XML.

Data available in SWDI are:

* Storm Cells from NEXRAD (Level-III Storm Structure Product)
* Hail Signatures from NEXRAD (Level-III Hail Product)
* Mesocyclone Signatures from NEXRAD (Level-III Meso Product)
* Digital Mesocyclone Detection Algorithm from NEXRAD (Level-III MDA Product)
* Tornado Signatures from NEXRAD (Level-III TVS Product)
* Preliminary Local Storm Reports from the NOAA National Weather Service
* Lightning Strikes from Vaisala NLDN

Find out more about SWDI at [http://www.ncdc.noaa.gov/swdi/#Intro](http://www.ncdc.noaa.gov/swdi/#Intro) and the API at [http://www.ncdc.noaa.gov/swdiws/](http://www.ncdc.noaa.gov/swdiws/).

********************

### Install `rnoaa`

Install and load `rnoaa` into the R session.

If you're on Windows, you may have to install Rtools. Run `devtools::has_devel()`. If you get a `TRUE`, you're okay. If not, [install Rtools](http://cran.r-project.org/bin/windows/Rtools/).

```{r install, eval=FALSE}
install.packages("devtools")
library(devtools)
ropensci::install_github("ropensci/rnoaa")
```

```{r load, message=FALSE, warning=FALSE}
library('rnoaa')
library('plyr')
```

### Search for nx3tvs data from 5 May 2006 to 6 May 2006

```{r}
swdi(dataset='nx3tvs', startdate='20060505', enddate='20060506')
```

### Get all 'nx3tvs' within 15 miles of latitude = 32.7 and longitude = -102.0

```{r}
out <- swdi(dataset='nx3tvs', startdate='20060506', enddate='20060507', radius=15, center=c(-102.0,32.7))
list(out$meta, head(out$data), head(out$shape))
```

### Use an id

```{r}
out <- swdi(dataset='warn', startdate='20060506', enddate='20060507', id=533623)
list(out$meta, head(out$data), head(out$shape))
```

### Get all 'plsr' within the bounding box (-91,30,-90,31)

```{r}
swdi(dataset='plsr', startdate='20060505', enddate='20060510', bbox=c(-91,30,-90,31))
```

### Get all 'nx3tvs' within the tile -102.1/32.6 (-102.15,32.55,-102.25,32.65)

```{r}
swdi(dataset='nx3tvs', startdate='20060506', enddate='20060507', tile=c(-102.12,32.62))
```

### Counts

Notes:

* stat='count' will only return metadata, nothing in the data or shape slots
* stat='tilesum:...' returns counts in the data slot for each date for that tile, and shape data
* Get number of 'nx3tvs' within 15 miles of latitude = 32.7 and longitude = -102.0

```{r}
swdi(dataset='nx3tvs', startdate='20060505', enddate='20060516', radius=15, center=c(-102.0,32.7), stat='count')
```

Get daily count nx3tvs features on .1 degree grid centered at `latitude = 32.7` and `longitude = -102.0`

```{r}
swdi(dataset='nx3tvs', startdate='20060505', enddate='20090516', stat='tilesum:-102.0,32.7')
```

### Get data in different formats

#### CSV format

```{r}
head(swdi(dataset='nx3tvs', startdate='20060505', enddate='20060506', format='csv')$data)
```

#### SHP format

```{r}
swdi(dataset='nx3tvs', startdate='20060505', enddate='20060506', format='shp', filepath='myfile')
```

#### KMZ format

```{r}
swdi(dataset='nx3tvs', startdate='20060505', enddate='20060506', format='kmz', radius=15, filepath='myfile.kmz')
```
