---
title: "NCDC workflow"
author: "Scott Chamberlain"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NCDC workflow}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo  =  FALSE, warning = FALSE, message = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse  =  TRUE,
  comment  =  "#>",
  purl  =  NOT_CRAN,
  eval  =  NOT_CRAN
)
```

This vignette is intended to demonstrate a workflow for using the NOAA NCDC data using the `ncdc*()` functions. It can be confusing to understand how to get at data you want - that's the motivation for this vignette. Other vignettes show more thorough and different examples for specific data sources.


## Load rnoaa

```{r load}
library('rnoaa')
```

## The workflow

* Look for weather stations & get station id(s)
* Find out what type of data is available for those stations
* Search for climate data for stations (optionally specify type of data to get)

### Look for weather stations & get station id(s)


```{r two, message=FALSE, warning=FALSE, cache=FALSE}
ids <- ncdc_stations(locationid='FIPS:12017')$data$id[1:13]
id <- "GHCND:US1FLCT0002"
```

Just information for one station

```{r}
ncdc_stations(stationid = id)
```


### Find out what type of data is available for those stations

There are various ways to look for data types available. First, __data categories__:

```{r}
ncdc_datacats(stationid = id)
```

Another way is looking for __data sets__:

```{r}
ncdc_datasets(stationid = id)
```

Yet another way is looking for __data types__:

```{r}
ncdc_datatypes(datasetid = "GHCND", stationid = id)
```

### Search for climate data for stations (optionally specify type of data to get)

Now that you know what kinds of data categories, data sets, and data types are available for your station you can search for data with any of those as filters.

Importantly, note that you have to specify three things in a call to the `ncdc` function:

* `datasetid`
* `startdate`
* `enddate`

Here, we are specifying the `datasetid`, `stationid`, `datatypeid`, `startdate`, and `enddate`

```{r}
ncdc(datasetid = "GHCND", stationid = id, datatypeid = "PRCP", startdate = "2012-10-01", enddate = "2013-01-01")
```
