---
title: "NOAA storms"
author: "Scott Chamberlain"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NOAA storms}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, echo  =  FALSE, warning = FALSE, message = FALSE}
NOT_CRAN <- identical(tolower(Sys.getenv("NOT_CRAN")), "true")
knitr::opts_chunk$set(
  collapse  =  TRUE,
  comment  =  "#>",
  purl  =  NOT_CRAN,
  eval  =  NOT_CRAN
)
```

Get NOAA wind storm tabular data, metadata, or shp files from International Best Track Archive for Climate Stewardship (IBTrACS).



Find out more about NOAA storms data at http://www.ncdc.noaa.gov/ibtracs/index.php?name=numbering

## Load rnoaa

```{r load, message=FALSE, warning=FALSE}
library('rnoaa')
library('plyr')
```

## Search for nx3tvs data from 5 May 2006 to 6 May 2006

```{r}
swdi(dataset='nx3tvs', startdate='20060505', enddate='20060506')
```

## Use an id

```{r}
out <- swdi(dataset='warn', startdate='20060506', enddate='20060507', id=533623)
list(out$meta, head(out$data), head(out$shape))
```

## Get all 'plsr' within the bounding box (-91,30,-90,31)

```{r}
swdi(dataset='plsr', startdate='20060505', enddate='20060510', bbox=c(-91,30,-90,31))
```

## Get all 'nx3tvs' within the tile -102.1/32.6 (-102.15,32.55,-102.25,32.65)

```{r}
swdi(dataset='nx3tvs', startdate='20060506', enddate='20060507', tile=c(-102.12,32.62))
```

## Counts

Notes:

* stat='count' will only return metadata, nothing in the data or shape slots
* stat='tilesum:...' returns counts in the data slot for each date for that tile, and shape data
* Get number of 'nx3tvs' within 15 miles of latitude = 32.7 and longitude = -102.0

Get daily count nx3tvs features on .1 degree grid centered at `latitude = 32.7` and `longitude = -102.0`

```{r}
swdi(dataset='nx3tvs', startdate='20060505', enddate='20090516', stat='tilesum:-102.0,32.7')
```

## Get data in different formats

### CSV format

```{r}
head(swdi(dataset='nx3tvs', startdate='20060505', enddate='20060506', format='csv')$data)
```

### SHP format

```{r}
swdi(dataset='nx3tvs', startdate='20060505', enddate='20060506', format='shp', filepath='myfile')
```

### KMZ format

```{r}
swdi(dataset='nx3tvs', startdate='20060505', enddate='20060506', format='kmz', radius=15, filepath='myfile.kmz')
```
