<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{rnoaa sparklines}
-->

rnoaa sparklines example
======

### Plot data from many stations

Load some libraries

```{r}
library(rnoaa)
library(scales)
library(lubridate)
library(maptools)
library(ggplot2)
library(doMC)
library(ggsubplot)
library(maps)
library(plyr)
library(stringr)
```

Find stations to get data from

```{r}
stations <- noaa_stations(datasetid = "GHCND", enddate = "2012-12-01", limit = 120)
res <- stations$data$id
```

Get data from stations. Note in the code below that we are using parallelization. You do not have to do this, you can simply set `.parallel=FALSE`, or equivalently, use `lapply` instead of `llply` (`llply` is from the `plyr` package).

```{r}
noaa_fw <- failwith(NULL, noaa)
registerDoMC(cores = 4)
dat <- llply(res, function(x) noaa_fw(datasetid = "GHCND", datatypeid = 'PRCP', stationid = x, startdate = '2010-06-01', enddate = '2010-09-30'), .parallel = TRUE)
dat <- compact(dat)
length(dat)
```

Make a `data.frame` and fix dates. 

```{r}
df <- ldply(dat, function(x) x$data)
df$date <- ymd(str_replace(as.character(df$date), "T00:00:00\\.000|T00:00:00", ""))
```

Get station lat and long data so that we can put data on a map.

```{r}
latlongs <- llply(res, function(x) 
  noaa_stations(x, datasetid = "GHCND")$data$meta[c("id", "latitude", "longitude")])
latlongs <- ldply(latlongs, function(x) as.data.frame(x))
df2 <- merge(df, latlongs, by.x = "station", by.y = "id")
head(df2)
```

Make a map

```{r}
world_map <- map_data("world")
p <- ggplot() + 
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), 
               fill = "white", color = "gray40", size = 0.2) + 
  annotate(geom = "text", x = -155, y = -55, 
           label = sprintf("Max value is\n %s mm", max(df2$value)/10))
p + 
  geom_subplot(aes(longitude, latitude, group = station, 
                   subplot = geom_line(aes(date, value)), size = 1), 
               ref = ref_vline(aes(fill = length(value)), thickness = 0.1), 
               width = rel(2), height = rel(5), data = df2) + 
  theme(legend.position = "none")
```