<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{rnoaa vignette}
-->

rnoaa vignette
======

### About the package

`rnoaa` is an R wrapper for the NOAA API.

********************

### Install `rnoaa`

Install and load `rnoaa` into the R session.

```{r install, eval=FALSE}
install.packages("devtools")
library(devtools)
install_github("rnoaa", "ropensci", ref="newapi")
```

```{r load, message=FALSE, warning=FALSE}
library(rnoaa)
library(plyr)
```

#### Get info on a station by specifcying a datasetid, locationid, and stationid

```{r two, message=FALSE, warning=FALSE, cache=FALSE}
noaa_stations(datasetid='GHCND', locationid='FIPS:12017', stationid='GHCND:USC00084289')
```

#### Search for data and get a data.frame or list

```{r three, message=FALSE, warning=FALSE, cache=FALSE}
out <- noaa(datasetid='NORMAL_DLY', stationid='GHCND:USW00014895', datatypeid='dly-tmax-normal')
```

##### See a data.frame

```{r four, message=FALSE, warning=FALSE, cache=FALSE}
out$data
```

#### Plot data, super simple, but it's a start

```{r six, message=FALSE, warning=FALSE, cache=FALSE}
out <- noaa(datasetid='NORMAL_DLY', stationid='GHCND:USW00014895', datatypeid='dly-tmax-normal')
noaa_plot(out)
```

### Plot data from many stations

#### Get table of all datasets
```{r seven, message=FALSE, warning=FALSE, cache=FALSE}
noaa_datasets()
```

#### Search for GHCND stations within 500 km of a lat/long point, take 10 of them

```{r eight, message=FALSE, warning=FALSE, cache=FALSE}
noaa_stations(datasetid='GHCND', locationid='FIPS:12017')
```

### Get data category data and metadata

```{r eleven, message=FALSE, warning=FALSE, cache=FALSE}
noaa_datacats(locationid='CITY:US390029')
```